---
title: "Final Report: Impact of Natural Birth Delivery and Vegetarian Diet on Improving Children’s Gut Microbiome Against Chronic Diseases
author: "Sanam Sajjadi"
date: "December 5, 2018"
output:github_document: default
bibliography: references.bib
csl: bioinformatics.csl
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Add about 2-3 pages here. Across the whole manuscript, you should cite at least 20 peer reviewed articles.







# Methods

## Study design

Add about half a page here. In this section instead of first person (I/we), use the authors of the paper you selected, since you'll just be describing what they did, based on the methods in their paper.

## Sample origin and sequencing

Add about half a page here. In this section instead of first person (I/we), use the authors of the paper you selected, since you'll just be describing what they did, based on the methods in their paper.

## Computational

These are the methods you used to do your bioinformatic analyses analyses. Should probably be between 0.5 and 1 pages. At a very minimum should include citations for DADA2 and phyloseq if you are doing an amplicon study, or other citations as appropriate.

# Results

## Subsections are ok in the results section too

```{r load-libraries, message = FALSE}
# Be sure to install these packages before running this script
# They can be installed either with the install.packages() function
# or with the 'Packages' pane in RStudio
# load general-use packages
library("dplyr")
library("tidyr")
library("knitr")
library("ggplot2")
```
```{r load-other-libraries}
# this package allows for the easy inclusion of literature citations in our Rmd
# more info here: https://github.com/crsh/citr
# and here:
# http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html
library("citr")
# These are the primary packages well use to clean and analyze the data
# this package needs to be installed from bioconductor -- it's not on CRAN
# see info here: https://benjjneb.github.io/dada2/dada-installation.html
library("dada2")
# This to export a fasta of our final denoised sequence variant
library("seqinr")
# To install this you have to install from GitHub
# See more info here: https://github.com/leffj/mctoolsr
# run this -- install.packages("devtools")
# and then this -- devtools::install_github("leffj/mctoolsr")
library("mctoolsr")
# And this to visualize our results
# it also needs to be installed from bioconductor
library("phyloseq")
# Be sure to load in the phyloseq data object created
load("output/phyloseq_obj.RData")
stopifnot(exists("phyloseq_obj"))
```
```{r load-packages-and-data}
# load packages as needed
library("dplyr")
load("output/phyloseq_obj.Rdata")
# load data produced from analysis scripts using
# something like load("output/processed_data.Rdata")
```



```{r make-mediterranean-diet-plot}
# We want to see the impact of diet
# On gut microbiome and as there are
# So many sequences in this data set
# Now we want to subset this data
# And look At gut microbiome of children with
# Mediterranean diet
# We look at the Phylum level
mediterranean_phyloseq_obj <- subset_samples(phyloseq_obj,
                              special_diet == "mediterranean diet")
plot_bar(mediterranean_phyloseq_obj, x = "Phylum")
```


**Figure1**: This figure is a plot bar that shows the different phyla of gut microbiome in children with Mediterranean diet.

Most of the bacteria were from Firmicutes phylum.


```{r make-vegeterian-diet-plot}
# We want to compare Mediterranean diet
# With vegeterian diet and
# AS we have a large data set
# We want to subset our dataset
# To look at gut microbiome of children
# With vegeterian diet
# Again we look at Phylum level
vegeterian_phyloseq_obj <- subset_samples(phyloseq_obj,
special_diet ==
"vegeterian diet (derived from sorghum, millet, black eyed pea)")
plot_bar(vegeterian_phyloseq_obj, x = "Phylum")
```

**Figure2**: This figure is a plot bar that shows the different phyla of bacteria in children with vegeterian diet.

Most of the sequences where from Bacteroidiates phylum.



```{r make-natural-childbirth-plot}
# Now we want to see the impcat of birth delivery method
# On Children's guts microbiome
# We subset our data set
# To look at gut microbiome of children
# That were born naturally
# We look at Phylum level
natural_phyloseq_obj <- subset_samples(phyloseq_obj,
birth_delivery == "natural childbirth")
plot_bar(natural_phyloseq_obj, x = "Phylum")
```


**Figure3**: This figure is a plot bar that shows the different phyla of bacteria in children that were born naturally.

Most of the bacteria were from Firmicutes and Bacteroidetes phyla.


```{r make-Cesarean-plot}
# now we want to look at microbiome of
# children's gut that were born with Cesarean
# As our data set is big, we make a new subset 
# For cesarean and then we look
# At different Phyla
cesarean_phyloseq_obj <- subset_samples(phyloseq_obj,
birth_delivery == "cesarean childbirth")
plot_bar(cesarean_phyloseq_obj, x = "Phylum")
```

**Figure4**: This figure is a plot bar that looks at different phyla of bacteria in children who were born with Cesarean section.

Most of the bacteria were from Firmicutes phylum.




```{r make-geographic-Burkina-plot}
# We want to see the impact of geographical location
# So we have to to subset our data for specific region
# Now we subset our big data
# To just look at Burkina Faso
# We look at Phylum
geographic_phyloseq_obj <- subset_samples(phyloseq_obj,
geographic_location_country ==
"Burkina Faso:Boulkiemde province,Boulpon")
plot_bar(geographic_phyloseq_obj, x = "Phylum")
```

**Figure5**: This figure is a plot bar that looks at different phyla of bacteria in children from Burkina Faso.

Most of the bacteria were from Bacteroidetes phylum.




```{r make-geographic-Florence-plot}
# After Burkina, now we want to look at 
# Gur bacteria in children from Florence
# So we have to subset our data to see different
# Phyla of bacteria in their gut
# We look at Phylum level
geographic_phyloseq_obj <- subset_samples(phyloseq_obj,
geographic_location_country == "Italy:Toscany, Florence")
plot_bar(geographic_phyloseq_obj, x = "Phylum")
```


**Figure6**: This figure is a plot bar that shows different phyla of bacteria in children from Florence.

Most of the bacteria were from Firmicutes phylum.






In addition to a minimum of 5-8 figures/tables (and associated captions), you should include sufficient text in this section to describe what your findings were. Remember that in the results section you just describe what you found, but you don't interpret it - that happens in the discussion. 2-3 pages.

# Discussion

Add around 3-4 pages interpreting your results and considering future directions one might take in analyzing these data.

# Sources Cited
